---
title: "Who Is the Greatest International Striker of All Time?"
subtitle: "A concise, data-led presentation"
author: "Dhruv Punia"
date: today
format:
  html:
    theme: default
    css: styles.css
    toc: false
---

<div class="quarto-navbar">
  <div class="navbar-content">
    <div class="logo-block">
      <span class="logo-tag">GOAT Striker</span>
    </div>
    <div class="nav-buttons">
      <a class="nav-button" href="index.html">🏠 Home</a>
      <a class="nav-button active" href="analysis.html">📊 Analysis</a>
      <a class="nav-button" href="visualizations.html">📈 Visualizations</a>
      <a class="nav-button" href="references.html">📚 References</a>
      <a class="nav-button" href="about.html">👤 About Me</a>
    </div>
  </div>
</div>

```{r}
#| echo: false
#| warning: false
#| message: false
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)

clean_goals <- read_csv("goalscorers_cleaned.csv", show_col_types = FALSE)
```

<main>
  <section class="section-card">
    <div class="section-header">
      <div>
        <h1 class="section-title">Analysis Dashboard</h1>
        <p class="section-intro">This report uses scoring volume, career longevity, and open-play quality to identify the strongest case for the greatest international striker.</p>
      </div>
      <div class="cta-row">
        <a class="cta-button" href="visualizations.html">Explore Visualizations</a>
        <a class="button-secondary" href="references.html">View References</a>
      </div>
    </div>

    <div class="section-description">Three premium analytics cards below present the core evidence in a clean data dashboard style.</div>
  </section>

  <section class="section-card">
    <div class="section-header">
      <div>
        <h2 class="section-title">Scoring Volume</h2>
        <p class="section-intro">Goals are the most direct measure of striker greatness. This view highlights the leading international scorers and shows who dominates purely by output.</p>
      </div>
    </div>

    ```{r}
    #| echo: false
    #| warning: false
    #| message: false
    ranked_scorers <- clean_goals |>
      filter(!is.na(scorer)) |>
      count(scorer, name = "total_goals", sort = TRUE) |>
      head(8)

    ggplot(ranked_scorers, aes(x = fct_reorder(scorer, total_goals), y = total_goals)) +
      geom_col(fill = "#4FC3F7", color = "#DCEBFF", width = 0.75) +
      geom_text(aes(label = total_goals), hjust = -0.18, size = 4, color = "#EAF5FF") +
      coord_flip() +
      theme_minimal(base_size = 14) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
      labs(title = NULL, x = NULL, y = "Total Goals") +
      theme(
        plot.background = element_rect(fill = "transparent", color = NA),
        panel.background = element_rect(fill = "transparent", color = NA),
        plot.title = element_text(face = "bold", size = 18, color = "#FFFFFF"),
        axis.text = element_text(color = "#E8EEF8"),
        axis.title = element_text(color = "#D8E2FF", face = "bold"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(color = "#1A2641"),
        legend.position = "none"
      )
    ```

    <div class="chart-caption">The top scorers are the first filter in the search for the greatest striker. High goal totals provide immediate evidence of elite attacking impact.</div>
  </section>

  <section class="section-card">
    <div class="section-header">
      <div>
        <h2 class="section-title">Longevity & Consistency</h2>
        <p class="section-intro">A durable career matters. This section explores how many years top international scorers remained active at the highest level.</p>
      </div>
    </div>

    ```{r}
    #| echo: false
    #| warning: false
    #| message: false
    longevity <- clean_goals |>
      filter(!is.na(scorer), !is.na(year)) |>
      group_by(scorer) |>
      summarise(
        first_year = min(year),
        last_year = max(year),
        years_active = last_year - first_year + 1,
        .groups = "drop"
      ) |>
      arrange(desc(years_active)) |>
      head(8)

    ggplot(longevity, aes(x = fct_reorder(scorer, years_active), y = years_active)) +
      geom_col(fill = "#8B5CF6", color = "#EAF5FF", width = 0.75) +
      geom_text(aes(label = years_active), hjust = -0.18, size = 4, color = "#EAF5FF") +
      coord_flip() +
      theme_minimal(base_size = 14) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
      labs(title = NULL, x = NULL, y = "Years Active") +
      theme(
        plot.background = element_rect(fill = "transparent", color = NA),
        panel.background = element_rect(fill = "transparent", color = NA),
        axis.text = element_text(color = "#E8EEF8"),
        axis.title = element_text(color = "#D8E2FF", face = "bold"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(color = "#1A2641"),
        legend.position = "none"
      )
    ```

    <div class="chart-caption">Longer careers signal sustained quality. Top scorers who remain productive over many years earn a stronger case for greatness.</div>
  </section>

  <section class="section-card">
    <div class="section-header">
      <div>
        <h2 class="section-title">Scoring Style</h2>
        <p class="section-intro">Open-play goals are a key indicator of a striker’s attacking versatility. This measure separates penalty-heavy totals from broader offensive impact.</p>
      </div>
    </div>

    ```{r}
    #| echo: false
    #| warning: false
    #| message: false
    top_scorers <- clean_goals |>
      filter(!is.na(scorer)) |>
      count(scorer, name = "total_goals", sort = TRUE) |>
      head(8)

    style_summary <- clean_goals |>
      filter(!is.na(scorer), !is.na(goal_type), scorer %in% top_scorers$scorer) |>
      count(scorer, goal_type, name = "goals") |>
      group_by(scorer) |>
      mutate(total_goals = sum(goals)) |>
      ungroup() |>
      filter(goal_type == "Open Play") |>
      mutate(open_play_share = round(100 * goals / total_goals, 1)) |>
      arrange(desc(open_play_share))

    ggplot(style_summary, aes(x = fct_reorder(scorer, open_play_share), y = open_play_share)) +
      geom_col(fill = "#4FC3F7", color = "#EAF5FF", width = 0.75) +
      geom_text(aes(label = paste0(open_play_share, "%")), hjust = -0.18, size = 4, color = "#EAF5FF") +
      coord_flip() +
      theme_minimal(base_size = 14) +
      scale_y_continuous(expand = expansion(mult = c(0, 0.12)), limits = c(0, 100)) +
      labs(title = NULL, x = NULL, y = "Open-Play Share (%)") +
      theme(
        plot.background = element_rect(fill = "transparent", color = NA),
        panel.background = element_rect(fill = "transparent", color = NA),
        axis.text = element_text(color = "#E8EEF8"),
        axis.title = element_text(color = "#D8E2FF", face = "bold"),
        panel.grid.major.y = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(color = "#1A2641"),
        legend.position = "none"
      )
    ```

    <div class="chart-caption">A high share of open-play goals suggests a striker is succeeding in active attacking situations, not only from penalties.</div>
  </section>
</main>

<footer class="footer">
  <div class="footer-inner">
    <span>GOAT Striker Analysis</span>
    <span>Built with Quarto + R</span>
    <span>© 2026</span>
  </div>
</footer>
