---
title: "Who Is the Greatest International Striker of All Time?"
subtitle: "Evidence, context, and a balanced GOAT model"
author: "Dhruv Punia"
date: today
format:
  html:
    theme: cosmo
    toc: true
runtime: shiny
---

```{r}
#| echo: false
#| warning: false
#| message: false
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)
library(shiny)

clean_goals <- read_csv("goalscorers_cleaned.csv", show_col_types = FALSE)

scorer_summary <- clean_goals |>
  filter(!is.na(scorer)) |>
  group_by(scorer, team) |>
  summarise(
    total_goals = n(),
    penalty_goals = sum(penalty, na.rm = TRUE),
    open_play_goals = sum(goal_type == "Open Play", na.rm = TRUE),
    active_years = n_distinct(year),
    first_year = min(year),
    last_year = max(year),
    .groups = "drop"
  ) |>
  mutate(
    open_play_share = round(100 * open_play_goals / total_goals, 1),
    penalty_share = round(100 * penalty_goals / total_goals, 1),
    goals_per_year = round(total_goals / active_years, 2),
    career_span = last_year - first_year + 1
  )

team_goals <- clean_goals |>
  group_by(team) |>
  summarise(team_goals = n(), .groups = "drop")

scorer_summary <- scorer_summary |>
  left_join(team_goals, by = "team") |>
  mutate(team_goal_share = round(100 * total_goals / team_goals, 1))

top_candidates <- scorer_summary |>
  arrange(desc(total_goals)) |>
  slice_head(n = 10)

decade_summary <- clean_goals |>
  filter(!is.na(decade)) |>
  group_by(decade) |>
  summarise(goals = n(), .groups = "drop") |>
  arrange(decade)

goal_type_by_scorer <- clean_goals |>
  filter(!is.na(scorer), !is.na(goal_type)) |>
  count(scorer, goal_type) |>
  group_by(scorer) |>
  mutate(total_goals = sum(n), share = n / total_goals) |>
  ungroup()

goat_model <- scorer_summary |>
  mutate(
    total_norm = (total_goals - min(total_goals)) / (max(total_goals) - min(total_goals)),
    open_play_norm = open_play_share / 100,
    efficiency_norm = (goals_per_year - min(goals_per_year)) / (max(goals_per_year) - min(goals_per_year)),
    model_score = 0.45 * total_norm + 0.30 * open_play_norm + 0.25 * efficiency_norm
  ) |>
  arrange(desc(model_score)) |>
  slice_head(n = 8)
```

# 1. Introduction / Hook

**Question answered:** Who really deserves the label “greatest international striker,” and why this is more than a goals leaderboard?

**What data is shown:** the dataset of every international goal, the leading scorers, and the metrics we will build from those records.

**Best visualization:** a short top-10 goals table plus a teaser statement that the raw leader is only the starting point.

**Insight:** The strongest case needs volume and context; raw totals are the first clue, not the final verdict.

The dataset records every international goal scored by an attacking player. In this investigation, we treat each goal as evidence and use it to test whether the all-time top names hold up once we measure quality, efficiency, and career profile.

```{r}
#| echo: false
#| results: asis
knitr::kable(
  top_candidates |>
    select(scorer, team, total_goals, active_years, open_play_share, penalty_share) |>
    arrange(desc(total_goals)) |>
    head(8),
  caption = "Top international scorers in the dataset, with goal type and career context."
)
```

# 2. Why total goals is not enough

**Question answered:** Does the top scorer also have the strongest claim to be the GOAT?

**What data is shown:** total goals for the leading scorers, plus note of penalty and open-play distribution.

**Best visualization:** a ranked bar chart of total goals for the top candidates.

**Insight:** Total goals expose the initial favorites, but they do not reveal how those goals were produced.

```{r}
#| echo: false
#| warning: false
#| message: false
ranked_goals <- scorer_summary |>
  arrange(desc(total_goals)) |>
  slice_head(n = 10)

ggplot(ranked_goals, aes(x = fct_reorder(scorer, total_goals), y = total_goals)) +
  geom_col(fill = "#1F4E79", width = 0.75) +
  geom_text(aes(label = total_goals), hjust = -0.1, size = 3, color = "#1F4E79") +
  coord_flip() +
  theme_minimal(base_size = 13) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "Top International Goalscorers",
    subtitle = "Raw goals are the first evidence, not the whole story",
    x = NULL,
    y = "Total Goals"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#4A4A4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

A professional analyst asks: how many of those goals were penalties? How many came from long careers? How many were scored in friendly fixtures? The dataset does not have every match context, so our next sections add metrics that make the case stronger.

# 3. Measuring scoring ability

**Question answered:** Which strikers earned their goals from open play rather than set-piece or penalty conversion?

**What data is shown:** goals by goal type for top candidates and the share of open-play scoring.

**Best visualization:** a stacked bar chart of goal-type distribution for the top scorers.

**Insight:** A high open-play share marks strikers who produce the kinds of goals that depend on movement, finishing, and chance creation.

```{r}
#| echo: false
#| warning: false
#| message: false
top_goal_type <- goal_type_by_scorer |>
  filter(scorer %in% ranked_goals$scorer[1:8]) |>
  mutate(goal_type = ifelse(goal_type == "Open Play", "Open Play", goal_type))

ggplot(top_goal_type, aes(x = fct_reorder(scorer, total_goals), y = share, fill = goal_type)) +
  geom_col(position = "fill") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Open Play" = "#C0392B", "Penalty" = "#F39C12", "Other" = "#7F8C8D")) +
  labs(
    title = "How Leading Scorers Built Their Goals",
    subtitle = "Open-play contribution versus penalties and other goal types",
    x = NULL,
    y = "Share of Goals",
    fill = "Goal Type"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

In a professional report, open-play scoring is treated as a quality filter. A striker with a high penalty share may still be great, but their profile is different from one whose goals come from active build-up and attacking movement.

# 4. Measuring efficiency

**Question answered:** Which striker produced the most goals per season of international scoring impact?

**What data is shown:** active years, total goals, and goals per active year for the leading candidates.

**Best visualization:** a scatter or bar comparison of goals-per-year versus total goals.

**Insight:** Efficiency exposes players who scored heavily in shorter windows and those who sustained output over decades.

```{r}
#| echo: false
#| warning: false
#| message: false
efficiency <- scorer_summary |>
  filter(total_goals >= 15) |>
  arrange(desc(goals_per_year)) |>
  slice_head(n = 12)

ggplot(efficiency, aes(x = fct_reorder(scorer, goals_per_year), y = goals_per_year, fill = total_goals)) +
  geom_col(width = 0.75) +
  coord_flip() +
  scale_fill_gradient(low = "#4E79A7", high = "#F28E2B") +
  labs(
    title = "Scoring Efficiency Among Leading Strikers",
    subtitle = "Goals per active international year is a proxy for striker productivity",
    x = NULL,
    y = "Goals per Active Year",
    fill = "Total Goals"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

A modern analysis gives weight to efficiency. A player who averages 5+ goals per international year is producing elite output, even if they played fewer seasons than the raw goals leader.

# 5. Comparing eras

**Question answered:** Do the scoring patterns change across decades, and how does that affect the GOAT conversation?

**What data is shown:** total international goals by decade and the evolution of major scoring periods.

**Best visualization:** a line chart or area chart of goals by decade.

**Insight:** Era context matters. A striker who dominated in one generation may not be directly comparable to someone from a later or earlier decade without adjustment.

```{r}
#| echo: false
#| warning: false
#| message: false
decade_summary <- decade_summary |>
  filter(!is.na(decade))

ggplot(decade_summary, aes(x = decade, y = goals)) +
  geom_line(color = "#2C3E50", size = 1.2) +
  geom_point(color = "#E74C3C", size = 3) +
  scale_x_continuous(breaks = decade_summary$decade) +
  theme_minimal(base_size = 13) +
  labs(
    title = "International Goal Volume by Decade",
    subtitle = "The dataset records changing scoring trends across eras",
    x = "Decade",
    y = "Goals"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

This chart reminds the audience that international scoring is not static. When we compare candidates, we are also comparing different calendars, match rhythms, and tournament schedules.

# 6. Build your own GOAT model

**Question answered:** How can you change the weights for volume, scoring quality, and efficiency to see who rises to the top?

**What you can do:** move the sliders below to set your preferred importance for total goals, open-play share, and goals per active year. The model will recalculate the ranking instantly.

```{r}
#| echo: false
#| warning: false
#| message: false

sliderInput("volume_weight", "Volume weight", min = 0, max = 1, value = 0.45, step = 0.05)
sliderInput("open_play_weight", "Open-play weight", min = 0, max = 1, value = 0.30, step = 0.05)
sliderInput("efficiency_weight", "Efficiency weight", min = 0, max = 1, value = 0.25, step = 0.05)
```

```{r}
#| context: server
#| echo: false
#| warning: false
#| message: false

ranking_reactive <- reactive({
  weights <- c(
    volume = input$volume_weight,
    open_play = input$open_play_weight,
    efficiency = input$efficiency_weight
  )

  total_weight <- sum(weights)
  if (total_weight == 0) {
    weights <- c(volume = 1/3, open_play = 1/3, efficiency = 1/3)
  } else {
    weights <- weights / total_weight
  }

  scorer_summary |>
    mutate(
      total_norm = (total_goals - min(total_goals)) / (max(total_goals) - min(total_goals)),
      open_play_norm = open_play_share / 100,
      efficiency_norm = (goals_per_year - min(goals_per_year)) / (max(goals_per_year) - min(goals_per_year)),
      model_score = weights["volume"] * total_norm +
        weights["open_play"] * open_play_norm +
        weights["efficiency"] * efficiency_norm
    ) |>
    arrange(desc(model_score)) |>
    mutate(model_score_pct = round(model_score * 100, 1))
})

output$top_player <- renderText({
  top_player <- ranking_reactive()$scorer[1]
  top_team <- ranking_reactive()$team[1]
  top_score <- ranking_reactive()$model_score_pct[1]
  paste0("Best player for your selected weights: ", top_player, " (", top_team, ") with a score of ", top_score, ".")
})

output$interactive_ranking <- renderTable({
  ranking_reactive() |>
    select(scorer, team, total_goals, open_play_share, goals_per_year, model_score_pct) |>
    head(8)
}, digits = 1)
```

```{r}
#| echo: false
verbatimTextOutput("top_player")
tableOutput("interactive_ranking")
```

This interactive version turns the model into a simple decision tool: the same dataset can now be evaluated from different perspectives depending on whether you value raw totals, open-play quality, or efficiency most.

# 7. Static GOAT ranking model

**Question answered:** How can we combine volume, scoring quality, and efficiency into a defensible composite score?

**What data is shown:** the model score for the top candidates and the components that feed it.

**Best visualization:** a horizontal ranking chart of model score.

**Insight:** A balanced model penalizes one-dimensional scorers and rewards those who bring sustained output and open-play quality.

```{r}
#| echo: false
#| warning: false
#| message: false
goat_ranking <- goat_model |>
  mutate(model_score = round(model_score * 100, 1))

ggplot(goat_ranking, aes(x = fct_reorder(scorer, model_score), y = model_score, fill = goals_per_year)) +
  geom_col(width = 0.75) +
  coord_flip() +
  scale_fill_gradient(low = "#A6CEE3", high = "#1F78B4") +
  labs(
    title = "Composite GOAT Score for Top Candidates",
    subtitle = "Volume, open-play quality, and efficiency combine into a final ranking",
    x = NULL,
    y = "GOAT Score",
    fill = "Goals per Year"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

```{r}
#| echo: false
#| results: asis
#| warning: false

knitr::kable(
  goat_ranking |>
    select(scorer, team, total_goals, active_years, open_play_share, goals_per_year, model_score),
  caption = "Top strikers ranked by the composite GOAT model."
)
```

This section is the culmination of the investigation: a model that does not simply reward the raw leader, but also gives credit to players who scored in open play and stayed productive across their international careers.

# 8. Final conclusion — the short answer

**Question answered:** Who do we pick as the best international striker based on this dataset and why?

**What you will see:** a short, chatty verdict and the compact evidence behind it.

**Best visualization:** the composite ranking chart above and a tiny summary table.

**Verdict (short):** below is the pick we recommend based on the composite score.

```{r}
#| echo: false
top_pick <- goat_ranking$scorer[1]
top_team <- goat_ranking$team[1]
top_score <- round(goat_ranking$model_score[1] * 100, 1)
cat(paste0("Our pick for the best international striker in this dataset: ", top_pick, " (", top_team, ") — model score: ", top_score, ".\n"))
```

**Why this pick?**

- It combines a high raw goal total with a strong open-play share and good efficiency across active years.
- The pick is not just the raw scorer — the model rewards balanced profiles.

If you want the raw leaderboard, the composite chart above shows how other top candidates compare.

## Limitations and next steps

- The dataset does not include minutes played or caps, so efficiency is proxied by goals per active year.
- We do not have full match importance or opponent strength for every goal.
- Adding minutes, competition type, and match result would let us refine the model (and maybe change the pick).

## About the Author

::: {.callout-note}
**Name:** [Dhruv Punia]  
**School:** [Central Jersey College Prep]  
**Grade:** [Rising 9th grader]

**Wh
y I chose this project:** I chose this project because it allowed me to explore a familiar sports question through data and visualization.

**What I learned:**
- How to clean and analyze a real dataset in R
- How to create clear charts that support a claim
- How to present findings in a concise Quarto report
:::

## References

Goalscorers dataset. (n.d.). Unpublished dataset used for this project.

R Core Team. (2024). R: A language and environment for statistical computing. R Foundation for Statistical Computing. https://www.R-project.org/

Wickham, H., et al. (2019). tidyverse: Easily install and load the tidyverse. https://CRAN.R-project.org/package=tidyverse

Wickham, H. (2016). ggplot2: Elegant graphics for data analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org/

Allaire, J. J., et al. (2024). Quarto. https://quarto.org/

Wickham, H., et al. (2023). dplyr: A grammar of data manipulation. https://CRAN.R-project.org/package=dplyr

Wickham, H., et al. (2023). readr: Read rectangular text data. https://CRAN.R-project.org/package=readr

Wickham, H. (2023). forcats: Tools for working with categorical variables. https://CRAN.R-project.org/package=forcats
