---
title: "Who Is the Greatest International Striker of All Time?"
subtitle: "A concise, data-led presentation"
author: "Dhruv Punia"
date: today
format:
  html:
    theme: cosmo
    toc: false
---

```{r}
#| echo: false
#| warning: false
#| message: false
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)

clean_goals <- read_csv("goalscorers_cleaned.csv", show_col_types = FALSE)
```

# Main Question

Who is the greatest international striker of all time?

This report uses three simple measures of greatness: scoring volume, longevity, and scoring style. Together, they build a clear and balanced case.

## 1. Scoring volume

The strongest starting point is total goals. A player who dominates this measure has the clearest initial case.

```{r}
#| echo: false
#| warning: false
#| message: false
ranked_scorers <- clean_goals |>
  filter(!is.na(scorer)) |>
  count(scorer, name = "total_goals", sort = TRUE) |>
  head(8)

ggplot(ranked_scorers, aes(x = fct_reorder(scorer, total_goals), y = total_goals)) +
  geom_col(fill = "#1F4E79", color = "white", width = 0.8) +
  geom_text(aes(label = total_goals), hjust = -0.15, size = 3.2, color = "#1F4E79") +
  coord_flip() +
  theme_minimal(base_size = 13) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Top International Goalscorers", x = NULL, y = "Total Goals") +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

- The leading player stands out clearly.
- This is the strongest first indicator of greatness because it measures direct output.

## 2. Longevity and consistency

Greatness is not only about a short peak. A long scoring career suggests durability and sustained quality.

```{r}
#| echo: false
#| warning: false
#| message: false
longevity <- clean_goals |>
  filter(!is.na(scorer), !is.na(year)) |>
  group_by(scorer) |>
  summarise(
    first_year = min(year),
    last_year = max(year),
    years_active = last_year - first_year + 1,
    .groups = "drop"
  ) |>
  arrange(desc(years_active)) |>
  head(8)

ggplot(longevity, aes(x = fct_reorder(scorer, years_active), y = years_active)) +
  geom_col(fill = "#2E8B57", color = "white", width = 0.8) +
  geom_text(aes(label = years_active), hjust = -0.15, size = 3.2, color = "#2E8B57") +
  coord_flip() +
  theme_minimal(base_size = 13) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(title = "Years Active for the Top Scorers", x = NULL, y = "Years Active") +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

- The strongest candidates also score over many seasons.
- Longevity strengthens the case because it shows sustained elite performance.

## 3. Scoring style

Open-play goals often reflect broader attacking quality than a penalty-heavy profile.

```{r}
#| echo: false
#| warning: false
#| message: false
top_scorers <- clean_goals |>
  filter(!is.na(scorer)) |>
  count(scorer, name = "total_goals", sort = TRUE) |>
  head(8)

style_summary <- clean_goals |>
  filter(!is.na(scorer), !is.na(goal_type), scorer %in% top_scorers$scorer) |>
  count(scorer, goal_type, name = "goals") |>
  group_by(scorer) |>
  mutate(total_goals = sum(goals)) |>
  ungroup() |>
  filter(goal_type == "Open Play") |>
  mutate(open_play_share = round(100 * goals / total_goals, 1)) |>
  arrange(desc(open_play_share))

ggplot(style_summary, aes(x = fct_reorder(scorer, open_play_share), y = open_play_share)) +
  geom_col(fill = "#C0392B", color = "white", width = 0.8) +
  geom_text(aes(label = paste0(open_play_share, "%")), hjust = -0.15, size = 3.2, color = "#C0392B") +
  coord_flip() +
  theme_minimal(base_size = 13) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)), limits = c(0, 100)) +
  labs(title = "Open-Play Share for Leading Scorers", x = NULL, y = "Open-Play Share (%)") +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

- The leading scorers are mostly built on open-play production.
- That makes the case for true striker quality stronger.

## Final takeaway

The best overall answer is the player who combines:

- the highest scoring total,
- the longest sustained career,
- and the strongest open-play profile.

That combination offers the clearest evidence for being the greatest international striker of all time.

## About the Author

::: {.callout-note}
**Name:** [Placeholder]  
**School:** [Placeholder]  
**Grade:** [Placeholder]

**Why I chose this project:** I chose this project because it allowed me to explore a familiar sports question through data and visualization.

**What I learned:**
- How to clean and analyze a real dataset in R
- How to create clear charts that support a claim
- How to present findings in a concise Quarto report
:::

## References

Goalscorers dataset. (n.d.). Unpublished dataset used for this project.

R Core Team. (2024). R: A language and environment for statistical computing. R Foundation for Statistical Computing. https://www.R-project.org/

Wickham, H., et al. (2019). tidyverse: Easily install and load the tidyverse. https://CRAN.R-project.org/package=tidyverse

Wickham, H. (2016). ggplot2: Elegant graphics for data analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org/

Allaire, J. J., et al. (2024). Quarto. https://quarto.org/

Wickham, H., et al. (2023). dplyr: A grammar of data manipulation. https://CRAN.R-project.org/package=dplyr

Wickham, H., et al. (2023). readr: Read rectangular text data. https://CRAN.R-project.org/package=readr

Wickham, H. (2023). forcats: Tools for working with categorical variables. https://CRAN.R-project.org/package=forcats
