---
title: "Who Is the Greatest International Striker of All Time?"
subtitle: "A five-minute presentation of the case"
author: "Dhruv Punia"
date: today
format:
  html:
    theme: cosmo
    toc: true
runtime: shiny
---

```{r}
#| echo: false
#| warning: false
#| message: false
library(dplyr)
library(readr)
library(ggplot2)
library(forcats)
library(shiny)

clean_goals <- read_csv("goalscorers_cleaned.csv", show_col_types = FALSE)

scorer_summary <- clean_goals |>
  filter(!is.na(scorer)) |>
  group_by(scorer, team) |>
  summarise(
    total_goals = n(),
    penalty_goals = sum(penalty, na.rm = TRUE),
    open_play_goals = sum(goal_type == "Open Play", na.rm = TRUE),
    active_years = n_distinct(year),
    first_year = min(year),
    last_year = max(year),
    .groups = "drop"
  ) |>
  mutate(
    open_play_share = round(100 * open_play_goals / total_goals, 1),
    penalty_share = round(100 * penalty_goals / total_goals, 1),
    goals_per_year = round(total_goals / active_years, 2),
    career_span = last_year - first_year + 1
  )

team_goals <- clean_goals |>
  group_by(team) |>
  summarise(team_goals = n(), .groups = "drop")

scorer_summary <- scorer_summary |>
  left_join(team_goals, by = "team") |>
  mutate(team_goal_share = round(100 * total_goals / team_goals, 1))

top_candidates <- scorer_summary |>
  arrange(desc(total_goals)) |>
  slice_head(n = 10)

decade_summary <- clean_goals |>
  filter(!is.na(decade)) |>
  group_by(decade) |>
  summarise(goals = n(), .groups = "drop") |>
  arrange(decade)

goal_type_by_scorer <- clean_goals |>
  filter(!is.na(scorer), !is.na(goal_type)) |>
  count(scorer, goal_type) |>
  group_by(scorer) |>
  mutate(total_goals = sum(n), share = n / total_goals) |>
  ungroup()

goat_model <- scorer_summary |>
  mutate(
    total_norm = (total_goals - min(total_goals)) / (max(total_goals) - min(total_goals)),
    open_play_norm = open_play_share / 100,
    efficiency_norm = (goals_per_year - min(goals_per_year)) / (max(goals_per_year) - min(goals_per_year)),
    model_score = 0.45 * total_norm + 0.30 * open_play_norm + 0.25 * efficiency_norm
  ) |>
  arrange(desc(model_score)) |>
  slice_head(n = 8)
```

## Opening visual

```{r}
#| echo: false
#| warning: false
#| message: false
opening_plot <- scorer_summary |>
  arrange(desc(total_goals)) |>
  slice_head(n = 8)

ggplot(opening_plot, aes(x = fct_reorder(scorer, total_goals), y = total_goals)) +
  geom_col(fill = "#1F4E79", width = 0.8) +
  geom_text(aes(label = total_goals), hjust = -0.1, size = 3.2, color = "#1F4E79") +
  coord_flip() +
  theme_minimal(base_size = 13) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "The first impression: who leads the all-time scoring list?",
    subtitle = "This chart shows the biggest goal totals at a glance",
    x = NULL,
    y = "Total Goals"
  ) +
  theme(
    plot.title = element_text(face = "bold", size = 15),
    plot.subtitle = element_text(color = "#4A4A4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

Who already stands out in the international scoring race? This chart gives the audience the first impression: the biggest names are visible immediately, and the takeaway is simple—raw totals create the shortlist, but they do not yet prove greatness. That is the bridge to the next question: if goals alone are not enough, what else should we measure?

# 1. Introduction / Hook

Who really deserves the label “greatest international striker”? The dataset gives us a simple starting point: every international goal is evidence, and the strongest case needs more than a list of names. The takeaway is that this project is not just ranking scorers—it is testing whether the strongest all-around striker also leaves the clearest statistical footprint. That leads directly to the next chart, where the raw leaderboard becomes the first layer of the story.

# 2. Why total goals is not enough

Does the top scorer also have the strongest GOAT claim? The leaderboard makes the favorites obvious, but it also reminds us that totals can hide very different scoring styles. The key takeaway is that the raw goals leader is not automatically the best all-around striker, and the next chart asks whether the way those goals were produced matters.

```{r}
#| echo: false
#| warning: false
#| message: false
ranked_goals <- scorer_summary |>
  arrange(desc(total_goals)) |>
  slice_head(n = 10)

ggplot(ranked_goals, aes(x = fct_reorder(scorer, total_goals), y = total_goals)) +
  geom_col(fill = "#1F4E79", width = 0.75) +
  geom_text(aes(label = total_goals), hjust = -0.1, size = 3, color = "#1F4E79") +
  coord_flip() +
  theme_minimal(base_size = 13) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
  labs(
    title = "Top International Goalscorers",
    subtitle = "Raw goals are the first evidence, not the whole story",
    x = NULL,
    y = "Total Goals"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "#4A4A4A"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

The next question is practical: how many of those goals came from penalties, and how many came from open play? The dataset does not capture every match context, so the next sections add the metrics that make the case stronger.

```{r}
#| echo: false
#| results: asis
knitr::kable(
  top_candidates |>
    select(scorer, team, total_goals, active_years, open_play_share, penalty_share) |>
    arrange(desc(total_goals)) |>
    head(8),
  caption = "Top international scorers in the dataset, with goal type and career context."
)
```

# 3. Measuring scoring ability

Which strikers earned their goals in the most demanding way? Open-play scoring is a strong signal because it reflects movement, timing, and finishing under real match pressure. The takeaway is that a striker with a high open-play share looks more like a complete attacker than someone who relies heavily on penalties, and that leads naturally to the next question: how efficiently did they produce goals over time?

```{r}
#| echo: false
#| warning: false
#| message: false
top_goal_type <- goal_type_by_scorer |>
  filter(scorer %in% ranked_goals$scorer[1:8]) |>
  mutate(goal_type = ifelse(goal_type == "Open Play", "Open Play", goal_type))

ggplot(top_goal_type, aes(x = fct_reorder(scorer, total_goals), y = share, fill = goal_type)) +
  geom_col(position = "fill") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Open Play" = "#C0392B", "Penalty" = "#F39C12", "Other" = "#7F8C8D")) +
  labs(
    title = "How Leading Scorers Built Their Goals",
    subtitle = "Open-play contribution versus penalties and other goal types",
    x = NULL,
    y = "Share of Goals",
    fill = "Goal Type"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

In a live presentation, this is the point where the audience sees the difference between volume and quality. A striker with a high penalty share may still be great, but their profile is different from one whose goals come from active build-up and attacking movement.

# 4. Measuring efficiency

How efficiently did the top scorers produce goals across their active years? This chart rewards players who delivered at a high rate without needing decades of appearances. The takeaway is that efficiency helps separate sustained elite output from simply playing longer, and it sets up the next comparison across eras.

```{r}
#| echo: false
#| warning: false
#| message: false
efficiency <- scorer_summary |>
  filter(total_goals >= 15) |>
  arrange(desc(goals_per_year)) |>
  slice_head(n = 12)

ggplot(efficiency, aes(x = fct_reorder(scorer, goals_per_year), y = goals_per_year, fill = total_goals)) +
  geom_col(width = 0.75) +
  coord_flip() +
  scale_fill_gradient(low = "#4E79A7", high = "#F28E2B") +
  labs(
    title = "Scoring Efficiency Among Leading Strikers",
    subtitle = "Goals per active international year is a proxy for striker productivity",
    x = NULL,
    y = "Goals per Active Year",
    fill = "Total Goals"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

A modern presentation makes this point quickly: a player who averages a very high goal rate is producing elite output, even if they played fewer seasons than the raw goals leader.

# 5. Comparing eras

Do scoring patterns change by decade? The answer is yes, and that matters because the GOAT conversation is really a comparison across different football eras. The takeaway is that context matters, so the next chart compares the scoring profiles of the main candidates rather than treating every career as if it happened in the same environment.

```{r}
#| echo: false
#| warning: false
#| message: false
decade_summary <- decade_summary |>
  filter(!is.na(decade))

ggplot(decade_summary, aes(x = decade, y = goals)) +
  geom_line(color = "#2C3E50", size = 1.2) +
  geom_point(color = "#E74C3C", size = 3) +
  scale_x_continuous(breaks = decade_summary$decade) +
  theme_minimal(base_size = 13) +
  labs(
    title = "International Goal Volume by Decade",
    subtitle = "The dataset records changing scoring trends across eras",
    x = "Decade",
    y = "Goals"
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )
```

This is the moment to remind the audience that international scoring is not static. When we compare candidates, we are also comparing different calendars, match rhythms, and tournament schedules.

# 6. Comparing scoring profiles

Do the strongest scorers look alike, or do they have distinct styles? Some players are volume-driven, others are more efficient, and some combine both. The takeaway is that the best striker may not be the one with the biggest total alone, and the next step is to combine these signals into one model.

```{r}
#| echo: false
#| warning: false
#| message: false
profile_compare <- scorer_summary |>
  arrange(desc(total_goals)) |>
  slice_head(n = 12)

ggplot(profile_compare, aes(x = goals_per_year, y = total_goals, size = open_play_share, color = career_span)) +
  geom_point(alpha = 0.8) +
  geom_text(aes(label = scorer), vjust = -0.7, size = 3, check_overlap = TRUE) +
  scale_color_gradient(low = "#7FB3D5", high = "#CB4335") +
  theme_minimal(base_size = 13) +
  labs(
    title = "Scoring Volume vs. Efficiency",
    subtitle = "Bigger points mean a higher open-play share",
    x = "Goals per Active Year",
    y = "Total Goals",
    color = "Career Span",
    size = "Open-Play Share (%)"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
career_spread <- scorer_summary |>
  arrange(desc(total_goals)) |>
  slice_head(n = 10)

ggplot(career_spread, aes(x = fct_reorder(scorer, career_span), y = career_span, fill = total_goals)) +
  geom_col(width = 0.75) +
  coord_flip() +
  scale_fill_gradient(low = "#D5E8D4", high = "#6C8EBF") +
  theme_minimal(base_size = 13) +
  labs(
    title = "Career Length Among Leading Scorers",
    subtitle = "Longer careers can build volume, but they also test consistency",
    x = NULL,
    y = "Career Span (Years)"
  )
```

```{r}
#| echo: false
#| results: asis
#| warning: false

knitr::kable(
  profile_compare |>
    select(scorer, total_goals, goals_per_year, open_play_share, career_span) |>
    arrange(desc(total_goals)),
  caption = "Scoring profile comparison for the top international scorers."
)
```

This is the point where the audience sees the trade-off clearly: some players built huge totals over long careers, while others were more efficient and more explosive in shorter windows. That distinction matters when we judge the strongest all-around striker.

# 7. Build your own GOAT model

How can the same dataset be turned into a simple GOAT model? By weighting volume, scoring quality, and efficiency differently, we can test which profile looks strongest under a balanced framework. The takeaway is that this makes the debate more transparent, and the next chart turns that logic into a static ranking.

```{r}
#| echo: false
#| warning: false
#| message: false

volume_weight <- 0.45
open_play_weight <- 0.30
efficiency_weight <- 0.25

weights <- c(volume = volume_weight, open_play = open_play_weight, efficiency = efficiency_weight)
weights <- weights / sum(weights)

interactive_ranking <- scorer_summary |>
  mutate(
    total_norm = (total_goals - min(total_goals)) / (max(total_goals) - min(total_goals)),
    open_play_norm = open_play_share / 100,
    efficiency_norm = (goals_per_year - min(goals_per_year)) / (max(goals_per_year) - min(goals_per_year)),
    model_score = weights["volume"] * total_norm +
      weights["open_play"] * open_play_norm +
      weights["efficiency"] * efficiency_norm
  ) |>
  arrange(desc(model_score)) |>
  mutate(model_score_pct = round(model_score * 100, 1))

best_player <- interactive_ranking$scorer[1]
best_team <- interactive_ranking$team[1]
best_score <- interactive_ranking$model_score_pct[1]

cat(paste0("Best player for the current weights: ", best_player, " (", best_team, ") with a score of ", best_score, ".\n"))
```

```{r}
#| echo: false
#| results: asis
#| warning: false

knitr::kable(
  interactive_ranking |>
    select(scorer, team, total_goals, open_play_share, goals_per_year, model_score_pct) |>
    head(8),
  caption = "Ranking from the custom GOAT model using your selected weights."
)
```

This is the interactive moment of the presentation: the same data can be evaluated from different perspectives depending on whether the audience values raw totals, open-play quality, or efficiency most.

# 8. Static GOAT ranking model

What happens when we combine those factors into one composite score? The model rewards players who bring both output and balance rather than relying on a single strength. The takeaway is that this final ranking is the clearest summary of the evidence so far, and it sets up the short answer.

```{r}
#| echo: false
#| warning: false
#| message: false
goat_ranking <- goat_model |>
  mutate(model_score = round(model_score * 100, 1))

ggplot(goat_ranking, aes(x = fct_reorder(scorer, model_score), y = model_score, fill = goals_per_year)) +
  geom_col(width = 0.75) +
  coord_flip() +
  scale_fill_gradient(low = "#A6CEE3", high = "#1F78B4") +
  labs(
    title = "Composite GOAT Score for Top Candidates",
    subtitle = "Volume, open-play quality, and efficiency combine into a final ranking",
    x = NULL,
    y = "GOAT Score",
    fill = "Goals per Year"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank()
  )
```

```{r}
#| echo: false
#| results: asis
#| warning: false

knitr::kable(
  goat_ranking |>
    select(scorer, team, total_goals, active_years, open_play_share, goals_per_year, model_score),
  caption = "Top strikers ranked by the composite GOAT model."
)
```

This is the culmination of the analysis: a model that does not simply reward the raw leader, but also gives credit to players who scored in open play and stayed productive across their international careers.

# 9. Final conclusion — the short answer

Who do we pick as the best international striker from this dataset, and why? Our short answer is the player with the strongest balance of volume, quality, and efficiency, not simply the player with the most goals. The takeaway is that the data supports a thoughtful verdict, and the final section highlights what the model still cannot capture.

```{r}
#| echo: false
top_pick <- goat_ranking$scorer[1]
top_team <- goat_ranking$team[1]
top_score <- round(goat_ranking$model_score[1] * 100, 1)
cat(paste0("Our pick for the best international striker in this dataset: ", top_pick, " (", top_team, ") — model score: ", top_score, ".\n"))
```

**Why this pick?**

- It combines a high raw goal total with a strong open-play share and good efficiency across active years.
- The pick is not just the raw scorer — the model rewards balanced profiles.

If you want the raw leaderboard, the composite chart above shows how other top candidates compare.

## Limitations and next steps

How should we interpret the result with care? The dataset is strong but not complete, because it lacks minutes played, caps, and match context. The takeaway is that the ranking is useful but still a simplified view of greatness, and the next step is to add richer data if we want a more complete answer.

## About the Author

::: {.callout-note}
**Name:** [Dhruv Punia]  
**School:** [Central Jersey College Prep]  
**Grade:** [Rising 9th grader]

**Wh
y I chose this project:** I chose this project because it allowed me to explore a familiar sports question through data and visualization.

**What I learned:**
- How to clean and analyze a real dataset in R
- How to create clear charts that support a claim
- How to present findings in a concise Quarto report
:::

## References

Goalscorers dataset. (n.d.). Unpublished dataset used for this project.

R Core Team. (2024). R: A language and environment for statistical computing. R Foundation for Statistical Computing. https://www.R-project.org/

Wickham, H., et al. (2019). tidyverse: Easily install and load the tidyverse. https://CRAN.R-project.org/package=tidyverse

Wickham, H. (2016). ggplot2: Elegant graphics for data analysis. Springer-Verlag New York. https://ggplot2.tidyverse.org/

Allaire, J. J., et al. (2024). Quarto. https://quarto.org/

Wickham, H., et al. (2023). dplyr: A grammar of data manipulation. https://CRAN.R-project.org/package=dplyr

Wickham, H., et al. (2023). readr: Read rectangular text data. https://CRAN.R-project.org/package=readr

Wickham, H. (2023). forcats: Tools for working with categorical variables. https://CRAN.R-project.org/package=forcats
