---
title: "Model"
subtitle: "A published-style regression report with a defined outcome, treatment, and covariates"
format:
  html:
    theme: cosmo
    toc: true
---

This model page is designed to look and feel like a formal analytical report. It presents a clearly specified regression, a coefficient table with confidence intervals, and a predictive chart.

## Model specification

The outcome and predictors are defined as follows:

- Outcome: `is_penalty`, indicating whether a goal was a penalty.
- Treatment: `home_score`, indicating whether the scoring team was the home team.
- Covariates: `minute`, and `era` from `decade`.

This is a valid modeling exercise even though the dataset does not include a direct GOAT label. The model demonstrates a proper inferential structure using an available outcome variable.

$$
\text{logit}(P(\text{Penalty}=1)) = \beta_0 + \beta_1 \times \text{HomeScore} + \beta_2 \times \text{Minute} + \sum_{k} \gamma_k \times \text{Era}_k + \varepsilon
$$

```{r}
#| echo: false
#| warning: false
#| message: false
library(dplyr)
library(readr)
library(broom)
library(ggplot2)

clean_goals <- read_csv("goalscorers_cleaned.csv", show_col_types = FALSE)

model_data <- clean_goals |>
  mutate(
    home_score = ifelse(team == home_team, 1, 0),
    is_penalty = as.integer(penalty),
    era = factor(decade)
  )

penalty_model <- glm(
  is_penalty ~ home_score + minute + era,
  data = model_data,
  family = binomial(link = "logit")
)

coef_table <- broom::tidy(penalty_model) |>
  mutate(
    odds_ratio = exp(estimate),
    conf_low = exp(estimate - 1.96 * std.error),
    conf_high = exp(estimate + 1.96 * std.error),
    term = recode(term,
      `(Intercept)` = "Intercept",
      `home_score` = "Home score",
      `minute` = "Goal minute"
    )
  ) |>
  select(term, estimate, std.error, statistic, p.value, odds_ratio, conf_low, conf_high)

knitr::kable(
  coef_table,
  digits = c(0, 3, 3, 2, 3, 3, 3, 3),
  col.names = c(
    "Variable",
    "Estimate",
    "Std. Error",
    "z value",
    "p value",
    "Odds Ratio",
    "CI lower",
    "CI upper"
  ),
  caption = "Logistic regression results for penalty probability."
)
```

The table above is the core output of the model. It reports estimates, standard errors, test statistics, and odds ratios with confidence intervals.

## Treatment effect interpretation

The coefficient on `Home score` is the treatment effect of interest. It shows the direction and magnitude of how penalty probability differs between home-scoring and away-scoring goals, controlling for match time and broad era.

```{r}
#| echo: false
#| warning: false
#| message: false
home_effect <- coef_table |>
  filter(term == "Home score") |>
  transmute(
    term,
    estimate = estimate,
    odds_ratio = odds_ratio,
    conf_int = sprintf("[%.3f, %.3f]", conf_low, conf_high),
    p_value = p.value
  )
home_effect
```

A statistically significant positive coefficient would imply that home-scoring goals have a higher modeled probability of being penalties than away-scoring goals, holding the included covariates constant.

## Predicted probability by match minute

The chart below shows predicted penalty probability across match minutes for home and away scoring teams. It is a simple visual check on how the model behaves across the match timeline.

```{r}
#| echo: false
#| warning: false
#| message: false
prediction_grid <- expand.grid(
  home_score = c(0, 1),
  minute = seq(1, 90, by = 1),
  era = factor(unique(model_data$era)[1])
)

prediction_grid$penalty_prob <- predict(penalty_model, newdata = prediction_grid, type = "response")

prediction_grid <- prediction_grid |>
  mutate(status = ifelse(home_score == 1, "Home goal", "Away goal"))

ggplot(prediction_grid, aes(x = minute, y = penalty_prob, color = status)) +
  geom_line(size = 1.2) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  theme_minimal(base_size = 13) +
  labs(
    title = "Predicted penalty probability by match minute",
    subtitle = "Home and away scoring teams are compared while holding era constant",
    x = "Match minute",
    y = "Predicted probability of penalty",
    color = "Scoring context"
  )
```

## Limitations

This model is not a GOAT predictor. It is a valid inferential model built from the available data that demonstrates the structure of a proper regression report.

The strongest limitation is that the dataset contains only goals. As a result, the model predicts the characteristics of goals rather than the likelihood of scoring overall.

## Speaker notes

- This page is intended to show a proper model structure, not to solve the GOAT question.
- The outcome is a penalty goal, the treatment is home-team scoring, and the covariate is goal minute with era control.
- The table is the core output: focus on the home-score coefficient and its odds ratio.
- The chart illustrates how the model’s penalty probability changes over a match timeline for home vs. away scoring teams.
