---
title: "Visualizations"
format:
  html:
    theme: default
    css: styles.css
    toc: false
---

<div class="quarto-navbar">
  <div class="navbar-content">
    <div class="logo-block">
      <span class="logo-tag">GOAT Striker</span>
    </div>
    <div class="nav-buttons">
      <a class="nav-button" href="index.html">🏠 Home</a>
      <a class="nav-button" href="analysis.html">📊 Analysis</a>
      <a class="nav-button active" href="visualizations.html">📈 Visualizations</a>
      <a class="nav-button" href="references.html">📚 References</a>
      <a class="nav-button" href="about.html">👤 About Me</a>
    </div>
  </div>
</div>

<main>
  <section class="section-card">
    <div class="section-header">
      <div>
        <h1 class="section-title">Visualizations</h1>
        <p class="section-intro">A polished gallery of charts that illustrate the dataset and highlight the top striker candidates in a professional dashboard layout.</p>
      </div>
      <div class="cta-row">
        <a class="cta-button" href="analysis.html">Back to Analysis</a>
      </div>
    </div>

    <div class="section-description">Beautiful visual storytelling is the foundation of this dashboard. Each chart is styled for clarity and matched to the galaxy theme for a cohesive portfolio experience.</div>
  </section>

  <section class="section-card">
    <div class="section-header">
      <div>
        <h2 class="section-title">Goals by Player</h2>
        <p class="section-intro">A modern bar chart showing the leading goalscorers in international play.</p>
      </div>
    </div>

    <div class="chart-frame">
      ```{r}
      #| echo: false
      #| warning: false
      #| message: false
      library(dplyr)
      library(readr)
      library(ggplot2)
      library(forcats)

      clean_goals <- read_csv("goalscorers_cleaned.csv", show_col_types = FALSE)
      ranked_scorers <- clean_goals |>
        filter(!is.na(scorer)) |>
        count(scorer, name = "total_goals", sort = TRUE) |>
        head(10)

      ggplot(ranked_scorers, aes(x = fct_reorder(scorer, total_goals), y = total_goals)) +
        geom_col(fill = "#8B5CF6", width = 0.75, alpha = 0.94) +
        geom_text(aes(label = total_goals), hjust = -0.16, size = 4, color = "#EAF5FF") +
        coord_flip() +
        theme_minimal(base_size = 14) +
        scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
        labs(title = NULL, x = NULL, y = "Goals") +
        theme(
          plot.background = element_rect(fill = "transparent", color = NA),
          panel.background = element_rect(fill = "transparent", color = NA),
          axis.text = element_text(color = "#E8EEF8"),
          axis.title = element_text(color = "#D8E2FF", face = "bold"),
          panel.grid.major.y = element_blank(),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_line(color = "#1A2641"),
          legend.position = "none"
        )
      ```
    </div>

    <p class="chart-caption">Top scorers are ranked by total goals, giving a clear view of the elite international strikers in the dataset.</p>
  </section>

  <section class="section-card">
    <div class="section-header">
      <div>
        <h2 class="section-title">Career Longevity</h2>
        <p class="section-intro">Compare how long the leading scorers remained active in international competition.</p>
      </div>
    </div>

    <div class="chart-frame">
      ```{r}
      #| echo: false
      #| warning: false
      #| message: false
      longevity <- clean_goals |>
        filter(!is.na(scorer), !is.na(year)) |>
        group_by(scorer) |>
        summarise(
          first_year = min(year),
          last_year = max(year),
          years_active = last_year - first_year + 1,
          .groups = "drop"
        ) |>
        arrange(desc(years_active)) |>
        head(10)

      ggplot(longevity, aes(x = fct_reorder(scorer, years_active), y = years_active)) +
        geom_col(fill = "#4FC3F7", width = 0.75, alpha = 0.94) +
        geom_text(aes(label = years_active), hjust = -0.18, size = 4, color = "#EAF5FF") +
        coord_flip() +
        theme_minimal(base_size = 14) +
        scale_y_continuous(expand = expansion(mult = c(0, 0.12))) +
        labs(title = NULL, x = NULL, y = "Years Active") +
        theme(
          plot.background = element_rect(fill = "transparent", color = NA),
          panel.background = element_rect(fill = "transparent", color = NA),
          axis.text = element_text(color = "#E8EEF8"),
          axis.title = element_text(color = "#D8E2FF", face = "bold"),
          panel.grid.major.y = element_blank(),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_line(color = "#1A2641"),
          legend.position = "none"
        )
      ```
    </div>

    <p class="chart-caption">Longevity helps distinguish strikers who stayed elite across many years, not just those with short peak performance.</p>
  </section>

  <section class="section-card">
    <div class="section-header">
      <div>
        <h2 class="section-title">Open Play Share</h2>
        <p class="section-intro">Show how much of each top scorer’s goals came from open play versus set pieces and penalties.</p>
      </div>
    </div>

    <div class="chart-frame">
      ```{r}
      #| echo: false
      #| warning: false
      #| message: false
      top_scorers <- clean_goals |>
        filter(!is.na(scorer)) |>
        count(scorer, name = "total_goals", sort = TRUE) |>
        head(10)

      style_summary <- clean_goals |>
        filter(!is.na(scorer), !is.na(goal_type), scorer %in% top_scorers$scorer) |>
        count(scorer, goal_type, name = "goals") |>
        group_by(scorer) |>
        mutate(total_goals = sum(goals)) |>
        ungroup() |>
        filter(goal_type == "Open Play") |>
        mutate(open_play_share = round(100 * goals / total_goals, 1)) |>
        arrange(desc(open_play_share))

      ggplot(style_summary, aes(x = fct_reorder(scorer, open_play_share), y = open_play_share)) +
        geom_col(fill = "#8B5CF6", width = 0.75, alpha = 0.94) +
        geom_text(aes(label = paste0(open_play_share, "%")), hjust = -0.18, size = 4, color = "#EAF5FF") +
        coord_flip() +
        theme_minimal(base_size = 14) +
        scale_y_continuous(expand = expansion(mult = c(0, 0.12)), limits = c(0, 100)) +
        labs(title = NULL, x = NULL, y = "Open-Play Share (%)") +
        theme(
          plot.background = element_rect(fill = "transparent", color = NA),
          panel.background = element_rect(fill = "transparent", color = NA),
          axis.text = element_text(color = "#E8EEF8"),
          axis.title = element_text(color = "#D8E2FF", face = "bold"),
          panel.grid.major.y = element_blank(),
          panel.grid.minor = element_blank(),
          panel.grid.major.x = element_line(color = "#1A2641"),
          legend.position = "none"
        )
      ```
    </div>

    <p class="chart-caption">A strong open-play share indicates a scorer who contributes in active attacking situations rather than relying on penalties.</p>
  </section>
</main>

<footer class="footer">
  <div class="footer-inner">
    <span>GOAT Striker Analysis</span>
    <span>Built with Quarto + R</span>
    <span>© 2026</span>
  </div>
</footer>
